---
title: "Stat 301 Project"
author: "Alyson Her"
date: "2025-04-08"
output: html_document
---
## Dataset
```{r}
library(tidyverse)
library(car)
library(readr)
Stat_Dataset <- read_csv("C:/Users/heral/Downloads/Spring 2025 Class/Analysis V  & M/Stat 301 Data project/Offical_Stat_Dataset.csv")
#Note: data is unbalanced
```


## Test 1: Does the average loan amount differs significantly across industries?
$H_0: ðœ‡_c=ðœ‡_a=ðœ‡_r=ðœ‡_n$
$H_a: at least 2 means are different 
$ðœ‡_c$- mean loan amount for Construction
$ðœ‡_a$- mean loan amount for Agriculture
$ðœ‡_r$- mean loan amount for Real estate
$ðœ‡_n$- mean loan amount for Recreation 
```{r}
names(Stat_Dataset)
Stat_Dataset

 
Avg_Industries<- Stat_Dataset %>% 
  select(NAICS_Names, DisbursementGross)
Avg_Industries


mod<- (aov(DisbursementGross~NAICS_Names, data= Avg_Industries)) 
summary(mod)


TukeyHSD(mod)
```
(F= 8.394, P-value= 0): Reject the null, there is enough evidence to suggest that the average loan amount differs significantly across industries.
Based on the Post HOC test between these industries Recreation-Construction, and Recreation-Real estate showed sig differences.


```{r}
Stat<- read_csv("C:/Users/heral/Downloads/Stat_Dataset (1).csv")
mod<- (aov(DisbursementGross~NAICS_Names, data= Stat)) 
summary(mod)
```


## Test 2: Is there a difference in loan terms and across industries and loan status?
Loan Status:
$H_0: ðœ‡_1=ðœ‡_2$
$H_a: ðœ‡_1\neqðœ‡_2$ 
$ðœ‡_1$- mean loan term based on PIF status
$ðœ‡_2$- mean loan term based on CHGOFF status
(F=163.4075, P=0): Reject the null. There is enough evidence to suggest that the mean loan term are sig different between PIF and CHGOFF loan status 

Industries: 
$H_0: ðœ‡_c=ðœ‡_a=ðœ‡_r=ðœ‡_n$
$H_a: at least 2 means are different 
$ðœ‡_c$- mean loan term for Construction
$ðœ‡_a$- mean loan term for Agriculture
$ðœ‡_r$- mean loan term for Real estate
$ðœ‡_n$- mean loan term for Recreation 
(F=4.36, P=0.005): Reject the null. There is enough evidence to suggest that that the mean loan term are sig different between industries 


Interaction (Loan Status : Industries):
H_0: There is no interaction between loan status and industries
H_a: There is an interaction between loan status and industries
(F=0.205, P=0.89): Fail to reject the null. There is no interaction between loan status and industry. The industry cannot predict the loan status. In the chart the the industries are fairly similar  
```{r}
anov<- lm(Term~MIS_Status+NAICS_Names+MIS_Status*NAICS_Names, data= Stat_Dataset)
Anova(anov, type="2")

model3.test <- aov(Term~NAICS_Names+MIS_Status, data = Stat_Dataset) # Which levels differ?
TukeyHSD(model3.test)
```
